---
title: "lab4"
author: "Qinwen Yang, AngelQinwen"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
biography <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(defense)
str(offense)
str(biography)
#Rename the Opponent_Opponent to Opponent in offense and defense.
defense <- defense %>% rename(Opponent = Opponent_Opponent)
offense <- offense %>% rename(Opponent = Opponent_Opponent)
```
Part One: Cleaning data

1.The player names and opponent names are factors.
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent <- as.factor(defense$Opponent)
str(defense)
offense$Name <- as.factor(offense$Name)
offense$Opponent <- as.factor(offense$Opponent)
str(offense)
biography$Name <- as.factor(biography$Name)
str(biography)


```
2.All offensive and defensive statistics are numerical. The Weight column in biography is also numeric.
```{r}
#?mutate_at
defense <- defense %>% mutate_at(vars(Tackles_Solo : Pass_PB), funs(as.numeric))
str(defense)
offense <- offense %>% mutate_at(vars(Receiving_REC : Passing_INT), funs(as.numeric))
str(offense)
biography <- biography %>% mutate(Weight = as.numeric(Weight))
str(biography)

```
3.Bonus: Change the Height column in biography into numeric. (Hint: Use a unit that works well)
```{r}
#First, separate 6-0 to two columns feet, inches
#Second, change two column types to numeric
#Third, apply centimeter formula

SepHeight <- biography %>% separate(Height, into=c('feet', 'inches'), sep = "-")
SepHeight$feet <- as.numeric(SepHeight$feet)
SepHeight$inches <- as.numeric(SepHeight$inches)
SepHeight <- SepHeight %>% mutate(Centimeter = SepHeight$feet * 30.48 + SepHeight$inches * 2.54)
SepHeight



```
4. At the end of this step you should have the following three clean data frames: defClean, offClean, and bioClean. Print the structures of these data frames.
```{r}
defClean <- defense
str(defClean)
offClean <- offense
str(offense)
bioClean <- biography
str(bioClean)
```

Part Two: Tidying (1)

1.Reformat the defClean data frame into a tidy format using gather. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.
```{r}
defClean <- defClean %>% gather(key=Type, value=Statistic, Tackles_Solo:Pass_PB)
defClean
```

2. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?
```{r}

ggplot(defClean, aes(x=Statistic))+geom_histogram(binwidth = 1)+facet_wrap(~Type) + ylab("Number of Players")+ggtitle("Relationship between Number of players and Statistic")


```

Conclusion: From the plot, it shows that Turnover_FR and Turnover_INT skills are most rare.


3. Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense.
```{r}
TwoState <- defense %>% 
  filter(Opponent == 'West Virginia'| Opponent == 'Kansas State') %>%
  select(Name, Opponent, Tackles_Solo) %>%
  spread(key=Opponent, value=Tackles_Solo)
TwoState 
TwoState <- TwoState %>% rename(KAN = `Kansas State`, West = `West Virginia`)
#ggplot(TwoState, aes(x=West, y=KAN))+geom_point(position = 'jitter')+geom_abline(slope=1, intercept = 0)
ggplot(TwoState, aes(x=West, y=KAN)) + geom_point(position = 'jitter')+xlab('solo tackles against West Virginia') + ylab('solo tackles against Kansas')+xlim(c(0,5))+ylim(c(0,10))


```

Conclusion: From this plot, since a large number of solo tackles means good defense, thus we are able to conclude that ISU have better defense against Kansas State.

Part Two: Tidying (2)

1. Separate the Hometown column into two columns, namely the city and the state. Print the first few rows of these two columns. (Hint: look at the sep= argument of separate)
```{r}
newHometown <- biography
newHometown <- newHometown %>% separate(Hometown, into = c('city', 'state'), sep=",") %>% select(city, state)
head(newHometown)
#table(newHometown$city)
#table(newHometown$state)


```
2. How many players are there from each state?
```{r}
NumPlayer <- biography %>% separate(Hometown, into=c('city', 'state'), sep = ",")  %>% select(Name, state)
head(NumPlayer)
(NumPlayer <- NumPlayer %>% group_by(state) %>% summarise(num_player = n()))


```

Part Three: Joining data frames

1. Find where do the offensive players come from by joining the offensive data and player biography data appropriately.
```{r}
newbiography <- biography %>% select(Name, Hometown)
newoffense <- offense %>% select(Name, Opponent, Receiving_REC)
inner_join(newbiography,newoffense)

```
2. Is there a relationship between the weight of a player and the receiving yards (Receiving_YDS)?
```{r}
biography$Name <- as.character(biography$Name) 
offense$Name <- as.character(offense$Name)
newoffense2 <- offense %>% group_by(Name) %>% summarise(Receiving_YDS = mean(Receiving_YDS))
joindata <- inner_join(biography,newoffense2, by='Name')
head(joindata)
ggplot(joindata, aes(x=Weight, y=Receiving_YDS))+geom_point(position = "jitter")+ggtitle("Relationship between the Weight and Receiving_YDS")
```

Conclusion:
There is not much an relationship between the weight of a player and the receiving yards, but in plot it shows an outlier which display the player Butler, Hakeem.

3. Explore more. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.

Question: I want to check if there is a relationship between the weight and the Tackles_Solo.
```{r}
biography$Name <- as.character(biography$Name) 
defense $Name <- as.character(defense$Name)
newdefense <- defense %>% group_by(Name) %>% summarise(Tackles_Solo = mean(Tackles_Solo))
joindata2 <- inner_join(biography, newdefense, by='Name')
head(joindata2)
ggplot(joindata2, aes(x=Weight, y=Tackles_Solo))+geom_point(position = "jitter")+ geom_abline(slope = 1, intercept = 0)+ ggtitle("Relationship between the Weight and Tackles_Solo")+xlim(c(175,300))+ylim(c(0,5))
  
  
```

Conclusion: There is not really clear and obvious relationship between Tackles_Solo and weight, there might be exits other factors that will effect Tackles_solo. In the plot, it also shows a outlier around weight 200lbs and the tackles_Solo is between 4 and 5.
