---
title: "Lab 4 team file"
author: "Jacob Shoemaker; jshoemaker2249, Qinwen Yang; AngelQinwen"
date: "3/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part one: Cleaning Data

1.Load data and rename columns
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
biography <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(defense)
str(offense)
str(biography)
#Rename the Opponent_Opponent to Opponent in offense and defense.
defense <- defense %>% rename(Opponent = Opponent_Opponent)
offense <- offense %>% rename(Opponent = Opponent_Opponent)
```

2. The player names and opponent names are factors.
```{r}

offense <- offense %>%
  mutate(Opponent = factor(Opponent),
         Name = factor(Name))

defense <- defense %>%
  mutate(Opponent = factor(Opponent),
         Name = factor(Name))

biography <- biography %>%
  mutate(Name = factor(Name))
```

3. All offensive and defensive statistics are numerical. The Weight column in biography is also numeric.
```{r}
#?mutate_at
defense <- defense %>% mutate_at(vars(Tackles_Solo : Pass_PB), funs(as.numeric))
offense <- offense %>% mutate_at(vars(Receiving_REC : Passing_INT), funs(as.numeric))
biography <- biography %>% mutate(Weight = as.numeric(Weight))


```
4. Bonus: Change the Height column in biography into numeric. 
```{r}
SepHeight <- biography %>% separate(Height, into=c('feet', 'inches'), sep = "-")
SepHeight
SepHeight$feet <- as.numeric(SepHeight$feet)
SepHeight$inches <- as.numeric(SepHeight$inches)
SepHeight <- SepHeight %>% mutate(Centimeter = SepHeight$feet * 30.48 + SepHeight$inches * 2.54)
SepHeight
```

5. At the end of this step you should have the following three clean data frames: defClean, offClean, and bioClean. Print the structures of these data frames.
```{r}
defClean <- defense
str(defClean)
offClean <- offense
str(offense)
bioClean <- biography
str(bioClean)
```

#Part Two: Tidying (1)

1 & 2.Reformat the defClean data frame into a tidy format using gather. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?
```{r}
defClean <- defClean %>% gather(key= Type, value = Statistic, 3:11)
defClean
ggplot(defClean, aes(x=Statistic)) +geom_histogram(bins=10) + facet_wrap(~Type) + ylab("Number of players")
```

From here, you can see that interceptions(Turnover_INT) and forced recoveries(Turnover_FR) are most rare. This makes sense, as turnovers are fairly infrequent in football, and for the part, cornerbacks and safeties are in the best position to intercept a pass.


3.Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense.
```{r}
defClean_Solo <- defClean[defClean$Type=='Tackles_Solo',] %>% group_by(Name)
complete <- filter(defClean_Solo, Opponent=='West Virginia' | Opponent=='Kansas State')%>% spread(key=Opponent, value=Statistic)
names(complete) <- c('Name', 'Type', 'KSU','WVU')
complete[is.na(complete)] <- 0
ggplot(complete, aes(x=WVU, y=KSU), group_by(Name)) + geom_point()+ xlim(c(0,5))+ylim(c(0,10)) +xlab("Tackles Against WVU") + ylab("Tackles Against KSU") + ggtitle("Solo Tackles Against Kansas State and West Virginia")

```


From this, better defense was played against Kansas State. While there are not many players that had a large number of solo tackles in both games, there was a larger clump of players that had between 1-4 solo tackles against KSU, that did not perform well against WVU. Therefore, those small increments of fewer tackles led to more overall solo tackles against Kansas State.


4. Separate the Hometown column into two columns, namely the city and the state. Print the first few rows of these two columns. (Hint: look at the sep= argument of separate)
```{r}
newHometown <- biography
newHometown <- newHometown %>% separate(Hometown, into = c('city', 'state'), sep=",") %>% select(city, state)
head(newHometown)
#table(newHometown$city)
#table(newHometown$state)
```

5. How many players are there from each state?
```{r}
NumPlayer <- biography %>% separate(Hometown, into=c('city', 'state'), sep = ",")  %>% select(Name, state)
head(NumPlayer)
(NumPlayer <- NumPlayer %>% group_by(state) %>% summarise(num_player = n()))
```

6. Find where do the offensive players come from by joining the offensive data and player biography data appropriately.
```{r}
newbiography <- biography %>% select(Name, Hometown)
newoffense <- offense %>% select(Name, Opponent, Receiving_REC)
inner_join(newbiography,newoffense)
```

7. Is there a relationship between the weight of a player and the receiving yards (Receiving_YDS)?
```{r}
biography$Name <- as.character(biography$Name) 
offense$Name <- as.character(offense$Name)
newoffense2 <- offense %>% group_by(Name) %>% summarise(Receiving_YDS = mean(Receiving_YDS))
joindata <- inner_join(biography,newoffense2, by='Name')
head(joindata)
ggplot(joindata, aes(x=Weight, y=Receiving_YDS))+geom_point(position = "jitter")+ggtitle("Relationship between the Weight and Receiving_YDS")
```

We would claim that there is not much of a trend between receiving yards and weight once a receiver weighs around 200lbs. The only receiver listed at 225 is Hakeem Butler, so it would make sense that there is an outlier for his performance.

8. Explore more. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.

Question 1:
```{r}
yrdsH <- left_join(offClean, SepHeight, by= 'Name')

yrdsH <- yrdsH[yrdsH$Receiving_YDS != 0,] %>% filter(!is.na(Name))
ggplot(yrdsH, aes(x= Centimeter, y = Receiving_YDS)) + geom_point() + ggtitle("Height vs. Receiving Yards") + xlab("Height") + ylab("Receiving Yards")
```

This surprised me a little to see that the tallest receivers were not the most successful in terms of overall yards. However, it is important to note that taller receivers may have more trouble if they get pushed around, as well as the fact that the Iowa State receiving corps is not all that deep.

Question 2: I want to check if there is a relationship between the weight and the Tackles_Solo.
```{r}
biography$Name <- as.character(biography$Name) 
defense $Name <- as.character(defense$Name)
newdefense <- defense %>% group_by(Name) %>% summarise(Tackles_Solo = mean(Tackles_Solo))
joindata2 <- inner_join(biography, newdefense, by='Name')
head(joindata2)
ggplot(joindata2, aes(x=Weight, y=Tackles_Solo))+geom_point(position = "jitter")+ geom_abline(slope = 1, intercept = 0)+ ggtitle("Relationship between the Weight and Tackles_Solo")+xlim(c(175,300))+ylim(c(0,5))
  
  
```


Conclusion: There is not really clear and obvious relationship between Tackles_Solo and weight, there might be exits other factors that will effect Tackles_solo. In the plot, it also shows a outlier around weight 200lbs and the tackles_Solo is between 4 and 5.